Spatio-temporal estimates of HIV risk group proportions for adolescent girls and young women across 13 priority countries in sub-Saharan Africa

### 2022-06-06 src/docs_paper/paper.Rmd
Close #118 by obtaining data from Juliana and computing the relevant statistics in the document

  **Results:**
  Data consisted of $291,465$ female survey respondents aged 15-29.
  Risk group proportions varied substantially across age groups (44.6\% of variation), countries (28.7\%), and between districts within each country (20.6\%), but changed little over time (1.6\%).
  Among women aged 20-29, we found a geographic delineation, north of which in eastern Africa, cohabiting (69.5\%) is more common than non-regular partner(s) (13.3\%), and south of which, non-regular partner(s) (52.6\%) are more common than cohabiting (27.9\%).
  Prioritisation based on behavioural risk, on top of location and age based prioritisation, reduced the proportion of population that must be reached to find half of all new infections from 20.1\% to 10.4\%.
  Female sex workers were 2.4\% of the at risk population but 15.9\% of all new infections.

@@

bibliography: citations.bib
output: rticles::plos_article
header-includes:
- \usepackage{tabularx}
# # plos.csl can be download and used locally
# csl: http://www.zotero.org/styles/plos
---

@@

In sub-Saharan Africa, adolescent girls and young women (AGYW) aged 15-29 are `r "num"`\% of the population but `r "num"`\% of new HIV infections [@unaids2021update].
HIV incidence amongst AGYW is `r "num"` times higher than among similarly aged males due to younger age at first sex, age patterns of sexual mixing, structural vulnerability and power imbalances, and increased susceptibility to HIV infection [@joint2016hiv].
AGYW have therefore been identified as a priority population for HIV primary prevention, and significant investments have been made in prevention programming [@saul2018dreams; @global2018measurement].

@@

@@

>>>>

  **Results:**
  Data consisted of `r "num"` female survey respondents aged 15-29.
  Risk group proportions varied substantially across age groups (`r "num"`\% of variation), countries (`r "num"`\%), and between districts within each country (`r "num"`\%), but changed little over time (`r "num"`\%).
  Among women aged 20-29, we found a geographic delineation, north of which in eastern Africa, cohabiting (`r "num"`\%) is more common than non-regular partner(s) (`r "num"`\%), and south of which, non-regular partner(s) (`r "num"`\%) are more common than cohabiting (`r "num"`\%).
  Prioritisation based on behavioural risk, in combiniation with location and age based prioritisation, reduced the proportion of population that must be reached to find half of all new infections from `r "num"`\% to `r "num"`\%.
  Female sex workers were `r "num"`\% of the at risk population but `r "num"`\% of all new infections.

@@

bibliography: citations.bib
output: rticles::plos_article
header-includes:
- \usepackage{tabularx}
# plos.csl can be download and used locally
csl: http://www.zotero.org/styles/plos
---

```{r echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)

priority_iso3 <- multi.utils::priority_iso3()
analysis_level <- multi.utils::analysis_level()
admin1_level <- multi.utils::admin1_level()

hiv_estimates <- readxl::read_excel(
  "HIV2021Estimates_Global_PLWHV-NI-Pop_6June2022.xlsx", 
  sheet = "Table"
) %>%
  filter(ISO3 %in% priority_iso3) %>%
  select(-c(`...1`, "UN", "ISO2", `...5`, "Region"))

population_comparison <- hiv_estimates %>%
  select(
    "ISO3",
    "L- Population by age and sex 15-19 ; Female", 
    "L- Population by age and sex 20-24 ; Female", 
    "L- Population by age and sex 25-29 ; Female",
    "L- Population aged 15-49 Male+Female"
  ) %>%
  mutate(pop_female_Y015_029 = `L- Population by age and sex 15-19 ; Female` +
    `L- Population by age and sex 20-24 ; Female` + 
    `L- Population by age and sex 25-29 ; Female`
  ) %>%
  summarise(
    total = sum(`L- Population aged 15-49 Male+Female`),
    subset = sum(pop_female_Y015_029)
  )

population_percent <- 100 * round(population_comparison$subset / population_comparison$total, digits = 2)

infections_comparison <- hiv_estimates %>%
  select(
    "ISO3",
    "L- New infections by age 15-19 ; Female", 
    "L- New infections by age 20-24 ; Female", 
    "L- New infections by age 25-29 ; Female",
    "L- New HIV infections (15-49) Male+Female"
  ) %>%
  mutate(inf_female_Y015_029 = `L- New infections by age 15-19 ; Female` +
    `L- New infections by age 20-24 ; Female` + 
    `L- New infections by age 25-29 ; Female`
  ) %>%
  summarise(
    total = sum(`L- New HIV infections (15-49) Male+Female`),
    subset = sum(inf_female_Y015_029)
  )

infections_percent <- 100 * round(infections_comparison$subset / infections_comparison$total, digits = 2)

inc_comparison <- hiv_estimates %>%
  select(
    "ISO3",
    "L- New infections by age 15-19 ; Female", 
    "L- New infections by age 20-24 ; Female", 
    "L- New infections by age 25-29 ; Female",
    "L- New infections by age 15-19 ; Male", 
    "L- New infections by age 20-24 ; Male", 
    "L- New infections by age 25-29 ; Male",
    "L- Population by age and sex 15-19 ; Female", 
    "L- Population by age and sex 20-24 ; Female", 
    "L- Population by age and sex 25-29 ; Female",
    "L- Population by age and sex 15-19 ; Male", 
    "L- Population by age and sex 20-24 ; Male", 
    "L- Population by age and sex 25-29 ; Male",
  ) %>%
  mutate(
    inf_female_Y015_029 = `L- New infections by age 15-19 ; Female` +
    `L- New infections by age 20-24 ; Female` + 
    `L- New infections by age 25-29 ; Female`,
    inf_male_Y015_029 = `L- New infections by age 15-19 ; Male` +
    `L- New infections by age 20-24 ; Male` + 
    `L- New infections by age 25-29 ; Male`,
    pop_female_Y015_029 = `L- Population by age and sex 15-19 ; Female` +
    `L- Population by age and sex 20-24 ; Female` + 
    `L- Population by age and sex 25-29 ; Female`,
    pop_male_Y015_029 = `L- Population by age and sex 15-19 ; Male` +
    `L- Population by age and sex 20-24 ; Male` + 
    `L- Population by age and sex 25-29 ; Male`
  ) %>%
  summarise(
    inc_female_Y015_029 = sum(inf_female_Y015_029) / sum(pop_female_Y015_029),
    inc_male_Y015_029 = sum(inf_male_Y015_029) / sum(pop_male_Y015_029)
  )

inc_ratio <- round(inc_comparison$inc_female_Y015_029 / inc_comparison$inc_male_Y015_029, digits = 1)
```

@@

In sub-Saharan Africa, adolescent girls and young women (AGYW) aged 15-29 are `r population_percent`\% of the population but `r infections_percent`\% of new HIV infections [@unaids2021update].
HIV incidence amongst AGYW is `r inc_ratio` times higher than among similarly aged males due to younger age at first sex, age patterns of sexual mixing, structural vulnerability and power imbalances, and increased susceptibility to HIV infection [@joint2016hiv].
AGYW have therefore been identified as a priority population for HIV primary prevention, and significant investments have been made in prevention programming [@saul2018dreams; @global2018measurement].

@@

## Supporting information

**S1**: Appendix
**S2**: Supplementary tables and figures

@@

# Tables

# Figures
